### 2021-12-13 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@62b5758e463482662104e6b7cbba81cb30dafd48 🚀

\fancypagestyle{normal}
{
  \fancyhf{}
  \fancyfoot[R]{\footnotesize\sffamily\thepage\ of \pageref*{LastPage}}
}
\begin{document}
\raggedright
\thispagestyle{empty}
{\Large\bfseries\sffamily Guidelines for the supervised learning of
species interactions}
\vskip 5em

@@

\begin{enumerate}
    \item The prediction of species interaction networks is gaining
momentum as a way to circumvent limitations in data volume. Yet,
ecological networks are challenging to predict because they are
typically small and sparse. Dealing with extreme class imbalance is a
challenge for most binary classifiers, and there are currently no
guidelines as to how predictive models can be trained.%
    \item Using simple mathematical arguments and numerical experiments
in which a variety of classifiers (for supervised learning) are trained
on simulated networks, we develop a series of guidelines related to the
choice of measures to use for model selection, and the degree of
unbiasing to apply to the training dataset.%
    \item Classifier accuracy and the ROC-AUC are not informative
measures for the performance of interaction prediction. PR-AUC is a
fairer assesment of performance. In some cases, even standard measures
can lead to selecting a more biased classifier because the effect of
connectance is strong. The amount of correction to apply to the training
dataset depends as a function of the classifier and the network
connectance.%
    \item These results reveal that training machines to predict
networks is a challenging task, and that in virtually all cases, the
composition of the training set needs to be experimented on before
performing the actual training. We discuss these consequences in the
context of the low volume of data.%
\end{enumerate}

@@

\begin{itemize}
\tightlist
\item
  example on diagnostic test: rare events are hard to detect even with
  really good models
\item
  summary of model challenges for networks
\item
  Strydom et al. (2021) importance of drawing on traits + validation is
  challenging + comparing across space
\end{itemize}

@@

Intro

@@

In this section, we will assume a network of connectance \(\rho\),
\emph{i.e.} having \(\rho S^2\) interactions (where \(S\) is the species
richness), and \((1-\rho) S^2\) non-interactions. Therefore, the vector
describing the \emph{true} state of the network is a column vector
\(\mathbf{o}^T = [\rho (1-\rho)]\) (we can safely drop the \(S^2\)
terms, as we will work on the confusion matrix, which ends up expressing
\emph{relative} values).

>>>>

\fancypagestyle{normal}
{
  \fancyhf{}
  \fancyfoot[R]{\footnotesize\sffamily\thepage\ of \pageref*{LastPage}}
}
\begin{document}
\raggedright
\thispagestyle{empty}
{\Large\bfseries\sffamily Validating machine learning predictions of
species interactions}
\vskip 5em

@@

\begin{enumerate}
    \item The prediction of species interactions is gaining momentum as
a way to circumvent limitations in data volume. Yet, ecological networks
are challenging to predict because they are typically small and sparse.
Dealing with extreme class imbalance is a challenge for most binary
classifiers, and there are currently no guidelines as to how predictive
models can be trained for this specific problem.%
    \item Using simple mathematical arguments and numerical experiments
in which a variety of classifiers (for supervised learning) are trained
on simulated networks, we develop a series of guidelines related to the
choice of measures to use for model selection, and the degree of
unbiasing to apply to the training dataset.%
    \item Classifier accuracy and the ROC-AUC are not informative
measures for the performance of interaction prediction. PR-AUC is a
fairer assesment of performance. In some cases, even standard measures
can lead to selecting a more biased classifier because the effect of
connectance is strong. The amount of correction to apply to the training
dataset depends on network connectance, on the measure to be optimized,
and only weakly on the classifier.%
    \item These results reveal that training machines to predict
networks is a challenging task, and that in virtually all cases, the
composition of the training set needs to be experimented on before
performing the actual training. We discuss these consequences in the
context of the low volume of data.%
\end{enumerate}

@@

The accuracy paradox is the basis of a number of problems in statistical
education, and lies in the fact that, when the desired class is rare, a
model that gets less and less performant will become more and more
accurate and useful, simply by (i) underpredicting true positive cases
and (ii) over-predicting false negatives. In other words, accuracy,
defined as the proportion of predictions that are correct, is often
useless as a measure of how predictive a model is. This is particularly
true in ecological networks; the desired class (presence of an
interaction between two species) is the one we care most about, and by
far the least commmon. Herein lies the core challenge of predicting
species interactions: the extreme imbalance between classes makes the
training of predictive models difficult. In a recent contribution,
Strydom et al. (2021) highlight that predictive models of interactions
can likely be improved by adding information (in the form of,
\emph{e.g.} traits), but that we do not have robust guidelines as to how
the predictive ability of these models should be evaluated, nor about
how the models should be trained. Here, by relying on simple derivations
and a series of simulations, we formulate a number of such guidelines,
specifically for the case of binary classifiers derived from thresholded
values.

@@

In this section, we will assume a network of connectance \(\rho\),
\emph{i.e.} having \(\rho S^2\) interactions (where \(S\) is the species
richness), and \((1-\rho) S^2\) non-interactions. Therefore, the vector
describing the \emph{true} state of the network is a column vector
\(\mathbf{o}^T = [\rho (1-\rho)]\) (we can safely drop the \(S^2\)
terms, as we will work on the confusion matrix, which ends up expressing
\emph{relative} values). We will apply skill and bias to this matrix,
and measure how a selection of performance metrics respond to changes in
these values, in order to assess their suitability for model evaluation.

@@
